Paper

### 2019-11-19 Paper_resubmit.Rmd
generating all new plots

@@

globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)

@@

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
```

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr,levels=c(1:22))) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = pg.samp %>% dplyr::select(Status, Prediction)
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
roc$model = 'CN Model'
roc$is = 'discovery'

@@

blues = RColorBrewer::brewer.pal(5,'Blues')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC*100))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

```{r}
pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) %>% dplyr::mutate(Pathology = factor(Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T))

sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))

@@

@@

setOgj <- function(nl) {
  if (nl >= 0 & nl < 3) return(1)
  if (nl >= 3 & nl < 7) return(2)
  if (nl >= 7) return(3)
}
pg.samp = pg.samp %>% mutate(ogj = factor(unlist(purrr::map(nl, setOgj)), ordered=T))

@@

ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )

# for (i in 1:nrow(sum.patient.data)) {
#   #print(i)
#   pt = sum.patient.data[[i, 'Patient']]
#   status = as.character(sum.patient.data[[i, 'Status']])
# 
#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
# 
#   fl = paste0('patient_', pt, '_',status,'.png')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
# }

pg.samp = pg.samp %>% dplyr::select(-ogj)
```

```{r, echo=F, fig.width=2, fig.height=4}
grid.arrange(p2$legend)

ggsave(filename='plots/spatial_mtn_pt20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
```

@@

rp = plot.risk.by.path(pg.samp)
rp
ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
```

@@

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path =arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)
```

@@

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
pg.samp = pg.samp %>% dplyr::mutate(nl = factor(nl, ordered=T), 
                                    Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), 
                                    Patient = factor(Patient, ordered=T), 
                                    Block = factor(Block, ordered=T), 
                                    Pathology = recode(Pathology, BE = 'NDBE'))

@@

dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
# Tiny heatmaps
plist = purrr::map(sum.patient.data$Patient, function(pt) {
  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block)

  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='') + theme(axis.text = element_blank(), plot.title = element_text(hjust = 0.5, face = 'bold', size=16))

  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance)), width=length(unique(patient$months.before.final)), units='in', dpi=300, limitsize=F)

  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block)

@@

p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
p11Loss = subset(p11, value < -1)
table(p11Loss$Status)

p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)

p11path['BE']/sum(p11path)
sum(p11path[c('HGD','IMC')])/sum(p11path)

table(p11Loss[c('Patient', 'Pathology')])

# Samples with a loss vs all Progressor samples
length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))

p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time

table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)

(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))

pi = patient.info
pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)

plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))

cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))

# So why might there be loss but no corresponding aberrant staining?
table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)

@@

## Top Features
```{r, echo=T, eval=F}
m = data.frame(matrix(nrow=nrow(features), ncol=0))

for (i in 1:nrow(features)) {
  feat = features[i,]

  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
  if(feat$end-feat$start > 5e6)
    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')

  segGL = subset(seg, value < -1)
  if(feat$coef > 0)
    segGL = subset(seg, value > 1)

  m[i, 'n.samples'] = length(unique(segGL$variable))
  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))

  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)

  m[i,'BE'] = segpath['BE']/sum(segpath)
  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
}
rownames(m) = features$label

range(m$ratio.samples.P)
summary(m$ratio.samples.P)
subset(m, ratio.samples.P == max(ratio.samples.P))
subset(features, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
#table(cdkGain[c('Patient', 'Pathology')])

cor.test(features$hazards, m$n.samples)

# Samples with a gain vs all samples

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
  filter(!grepl('^X|Y',`Genome Location`)) %>% 
  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)

genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) & `Mutation Types` != 'T') %>% 
  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
  dplyr::mutate(value = min(globalCNdf$value), Patient = '')

#genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
#chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)

#ov = findOverlaps(chGR, genesGR)
#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
#colnames(genes)[1] = c('chr')

g17 = 
  plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
  labs(title="All Samples") + theme(legend.position = 'none')
g17

g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)

```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
chGR$Genes = ''
for (hit in unique(queryHits(ov))) {
  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
}

for (chr in unique(chh$chr)) {  
  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
  print(p)
}

annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2

p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5 ) + theme(legend.position = 'none')

@@

(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
```

@@

#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value))

@@

dir.create('plots/validation/heatmaps/P',recursive = T, showWarnings = F)
dir.create('plots/validation/heatmaps/NP',recursive = T, showWarnings = F)

>>>>

patient.info = patient.info %>% mutate( Pathology = recode(Pathology, 'BE'='NDBE'))

@@

globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')

gp = globalCN.plot + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)

gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)

@@

cvRR = features

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
```

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr,levels=c(1:22))) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = pg.samp %>% dplyr::select(Status, Prediction)
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)

roc$model = 'CN Model'
roc$is = 'discovery'

@@

blues = RColorBrewer::brewer.pal(5,'Blues')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_y_continuous(limits=c(0,100)) +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3, height=4, units="in", limitsize=F, dpi = 300)

@@

```{r}
pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))

@@

mini = pcal + labs(x='P(P)', y='P:NP', title='') + theme(legend.position = 'none') + 
  scale_x_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = cuts, labels = c(0,'','',0.3,'',0.5,'','','','',1)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = seq(0,1,0.25), labels = seq(0,1,0.25)) + 
  theme(axis.text = element_text(size=14))
ggsave(filename='plots/pred_calib_mini.png', plot=mini, width=4, height=4, units='in', dpi=300)

@@

setOgj <- function(nl) {
  if (nl >= 0 & nl < 2) return(1)
  if (nl >= 2 & nl < 4) return(2)
  if (nl >= 4 & nl < 7) return(3)
  if (nl >= 7) return(4)
}
pg.samp = pg.samp %>% mutate(ogj = factor(unlist(purrr::map(nl, setOgj)), ordered=T))

@@

ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p2$grob), width=10.5, height=4.5, units='in', dpi=300)
ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p6$grob), width=10.5, height=4.5, units='in', dpi=300)

grid.arrange(p2$grob, p6$grob, ncol=1)

# for (i in 1:nrow(sum.patient.data)) {
#   print(i)
#   pt = sum.patient.data[[i, 'Patient']]
#   status = as.character(sum.patient.data[[i, 'Status']])
# 
#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
# 
#   fl = paste0('patient_', pt, '_',status,'.png')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
# }

pg.samp = pg.samp %>% dplyr::select(-ogj)

ggsave('plots/spatial/legend.png', plot=grid.arrange(p2$legend), width=2,height=2, units='in', dpi=300)
```

@@

rp = plot.risk.by.path(pg.samp)
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

@@

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path =arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), nudge_y=1) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + 
  theme(legend.position = 'bottom', text = element_text(size=14))
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
```

@@

```{r endos, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
pg.samp = pg.samp %>% dplyr::mutate(nl = factor(nl, ordered=T), 
                                    Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), 
                                    Patient = factor(Patient, ordered=T), 
                                    Block = factor(Block, ordered=T))

@@

dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
# Tiny heatmaps
plist = purrr::map(sum.patient.data$Patient, function(pt) {
  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup

  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))

  ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
  wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )

  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)

  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup

@@

p11 = filter(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
p11Loss = p11 %>% filter(value < -1)

p11Loss %>% filter(value < -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% full_join(
  p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
  kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)

p11Loss %>% group_by(Status, Hospital.Research.ID) %>% tally

p11Path = patient.info %>% filter(Samplename %in% p11Loss$variable) %>% dplyr::group_by(Pathology) %>% tally(name='samples')

p11Path %>% dplyr::mutate('ratio'=round(samples/sum(p11Path$samples),2)) %>% 
kable(caption='Ratio of all Chr17 p11 loss samples by pathology' ) %>% kable_styling(full_width = F)
```

Samples with a loss vs all Progressor samples
Ratio of samples with p11 loss vs all progressor samples: `r length(unique(p11Loss$variable))/length(unique(filter(globalCNdf, Status == 'Progressor')$variable))`

Ratio of progressor patients with a p11 loss in at least 1 sample: `r length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)`

```{r}
patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 status for samples with p11 loss') %>% kable_styling(full_width = F)

pi = patient.info
pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
#plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))

cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
```

## Top Features

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
  filter(!grepl('^X|Y',`Genome Location`)) %>% 
  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
  dplyr::filter(start != '' | end != '') %>%
  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)

genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) | `Mutation Types` != 'T') %>% 
  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
  dplyr::mutate(value = min(globalCNdf$value), Patient = '', chr = factor(chr,levels=levels(chr.lengths$chr),ordered=T), start = as.numeric(start), end=as.numeric(end)) %>% arrange(chr, start)

genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
chGR = makeGRangesFromDataFrame(features, keep.extra.columns = T)

#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
#colnames(genes)[1] = c('chr')

g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))

ggsave(filename='plots/hazards/chr17.png', plot=g17, width=8, height=4, units='in', dpi=300)

@@

```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
ov = findOverlaps(chGR, genesGR)
chGR$Genes = ''
for (hit in unique(queryHits(ov))) {
  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
}

for (chr in unique(features$chr)) {  
  print(chr)
  p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
    scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
    theme(strip.text.x = element_blank(), text=element_text(size=14)) + labs(title=paste('Chromosome',chr))
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=8, height=4, units='in', dpi=300)
}

annoG = ccgenes %>% filter(grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', `Gene Symbol`)) %>% arrange(chr) %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  mutate(ymin = round(range(globalCN.plot$data$mean.value))[1]/2, ymax = round(range(globalCN.plot$data$mean.value))[2]/2) %>%
  mutate(chr = factor(chr, levels=levels(chr.lengths$chr)), start = as.numeric(start), end = as.numeric(end))

p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=`Gene Symbol`), color='red', size=1.5 ) + theme(legend.position = 'none')

@@

#(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
```

@@

#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value)) + ylim(-30,20)

@@

dir.create('plots/validation/heatmaps/P/tiny',recursive = T, showWarnings = F)
dir.create('plots/validation/heatmaps/NP/tiny',recursive = T, showWarnings = F)
